---
title: "quandle 1 nse"
author: "Vijay Baby Joseph"
date: "27 August 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Quandl)
library(dplyr)
library(quantmod)
library(tseries)
library(forecast)
library(xts)
library(ggplot2)
library(data.table)
```


#A little about the Qnifty...

The Quantum Nifty ETF (QNF) is a replica of the NIFTY 50 - the stock index of NSE. The stocks that form a part of NIFTY 50 also form a part of QNF`s portfolio in the same proportion and weightage as that of the NIFTY 50.

This analysis will focus on Qnifty and its relation to the Nifty with the aim to forcast future performance
using basic arima, autoarima and ets.

#Loading the data and checking the dataset
```{r cars}

#Load the data

Quandl.api_key('cmgJvqW3MRmXc8M5RHzU')

#data1=read.csv("C:/Users/Administrator/Desktop/DataScience/nse index/NSE-LT.csv")
data = Quandl("NSE/QNIFTY", start_date="2012-01-01", end_date="2017-12-31")
test = Quandl("NSE/QNIFTY", start_date="2018-01-01", end_date="2018-08-29")

niftyone=Quandl("NSE/NIFTY_50", start_date="2012-01-01", end_date="2017-12-31")
niftytest = Quandl("NSE/NIFTY_50", start_date="2018-01-01", end_date="2018-08-29")


# checking the dataset
head(data)
str(data)
class(data)

#avg trading days in a year
nrow(data)/5
```

#Moving Average, Volatility, Start, End , etc. 

```{r}

#plotting moving average and volatility
rolling_average <- rollapply(data[,4],200,mean)
plot.ts(rolling_average)

rolling_volatility <- rollapply(data[,4],200,sd)
plot.ts(rolling_volatility)



#choosing cols and converting to time series

tsdata=data%>%select("Date","Close")

#stocks=ts(rev(tsdata$Close),start=c(2012, 1),frequency=365)
stocks<- xts(tsdata[,-1],order.by=as.Date(tsdata[,1], "%Y-%m-%d"))# xts takes daily frequency by default

head(stocks)

#mode(stocks)

# structure
str(stocks)

#start
start(stocks)

#end
end(stocks)

#frequency
frequency(stocks)

#cycle(stocks)

#time series summary
summary(stocks)


#daily returns
returns <- diff(stocks, arithmetic=FALSE ) - 1
head(returns)

## simlple line chart of Qnifty ETF
chartSeries(stocks, name='Qnifty')


#merging plot data and deleting na values caused by missing trading values from source

qnf = data.frame(stocks)
qnf$Date = as.Date(rownames(qnf))
head(qnf)
nft=niftyone%>%select('Date','Close')


setDT(nft)
setDT(qnf)

plot0 = nft[qnf, on = c('Date','Date')]
plot0=na.omit(plot0)
plot0=plot0%>%select(-Date.1)


range01 <- function(x){(x-min(x))/(max(x)-min(x))}


rangenifty=range01(plot1$NiftyIndex)
plot0$rangeqnf=range01(as.numeric(plot0$stocks))
plot0$rangenft=range01(as.numeric(plot0$Close))


# Plotting both Qnifty and Nifty on a normalised graph

ggplot(plot0, aes(Date)) + 
  geom_line(aes(y = plot0$rangeqnf, colour = "QNifty Price")) + 
  geom_line(aes(y = plot0$rangenft, colour = "Nifty Index"))+ggtitle("Qnifty vs Nifty (2012-2017)")+ xlab("Date") + ylab("Normalized Price/Index")

```


#Using decompose() to split the ts into seasonality, trend and error components.
```{r}

#Conversion to ts and plotting

attr(stocks, 'frequency') <- 224  # Set the frequency of the xts object to no. of annaual trading days
periodicity(stocks)             # check periodicity 
plot(decompose(as.ts(stocks)))  # Decompose after conversion to ts

decomp <- decompose(as.ts(stocks) ) # type="mult" use type = "additive" for additive components
plot (decomp)


#Dickey-Fuller Test
adf.test(as.ts(stocks)) # p-value >0.05 indicates the TS is non stationary


```

#Using forecast::stl() to split the ts into seasonality, trend and error components and making it stationary
```{r}

ts.stl <- stl(stocks,"periodic")  # decompose the TS
plot(ts.stl)

ts.sa <- seasadj(ts.stl)  # de-seasonalize

plot(stocks, type="l")  # original 
plot(ts.sa, type="l")  # adjusted

# seasonal frequency set as 224  for annual average trading days - 5 years prices preseneted as diff colors
seasonplot(ts.sa, 224, col=rainbow(5), year.labels=TRUE, main="QNifty") 

# Make it stationary
ndiffs(as.ts(stocks))  # number of differences need to make it stationary = 1

stationaryTS <- diff(stocks, differences= 1)
plot(stationaryTS, type="l", main="Differenced and Stationary")  # appears to be stationary

```


#Autocorrelation and Partial-Autocorrelation:

The final objective is to find pdq
AR I MA
p  d  q

-p is the number of autoregressive terms

-d is the number of nonseasonal differences needed for stationarity - or the no. of differentiations required 
in this case one as shown in the ndiffs function

-q is the number of lagged forecast errors

```{r}
# both acf() and pacf() generates plots by default
acf(as.ts(stocks)) # autocorrelation - determines q value for MA in arima
pacf(as.ts(stocks))  # partial autocorrelation - determines p value in arima

#its a non stationary series as as the acf falls very gradually and the adf test above gives a p val greter than 0.05


#Differencing subtracts each data point in the series from its successor to make it stationary.
plot(diff(log(as.ts(stocks))))

acf(diff(log(as.ts(stocks)))) # value of q is 1
pacf(diff(log(as.ts(stocks)))) # value of p is 0

# differentiation is used once so d=1 - this is also whown in the ndiff function above
```

#ARIMA 
```{r}

#arima model 
arimamodel=arima(log(as.ts(stocks)), c(0,1,1), seasonal=list(order=c(0,1,1), period=224))

#predict
pred=predict(arimamodel, n.ahead = 109)

corpred=2.718^pred$pred
corpred=round(corpred,digits=2)

plotone=data.frame(test$Date,corpred,test$Close)

#plotting the prices for the next 109 trading days - from jan 2018 till end of august 2018
ggplot(plotone, aes(plotone$test.Date)) + 
  geom_line(aes(y = plotone$corpred, colour = "QNifty Price")) + 
  ggtitle("Qnifty (Jan - Aug 2018) Price Prediction")+ xlab("Date") + ylab("Price")


#compare with test
corpred #predicted prices
test$Close #actual prices 


#Percentage Error
names(plotone)<-c("Date", "Actual Index","Forecasted Price")

perc_error= function(act,fcst){
  pe=((act-fcst)/(act))
  mpe=mean(pe)
return(mpe)}


perc_error(plotone$`Actual Index`,plotone$`Forecasted Price`)
```





#AUTOMATED FORECASTING USING ARIMA AND ETS
```{r}
#auto.arima model

logstock=log(stocks)
fit<-auto.arima(as.ts(logstock),D=1)
fit


#ETS model
fit2 <- ets(as.ts(stocks),model = "ZZZ",damped = FALSE, lambda = TRUE)
fit2


#Forecasted Values from auto.arima
forecastedval=forecast(fit,h=109)

forecastedval=as.numeric(forecastedval$mean)
corforcastedval= exp(forecastedval)
corforcastedval=round(corforcastedval,digits=4)
corforcastedval

ts.plot(corforcastedval, gpars=list(xlab="Days-109", ylab="Predicted Price"))

#Percentage Error

corforcastedval=as.data.frame(corforcastedval)
test$Close=as.data.frame(test$Close)

df<-data.frame(test$Close,corforcastedval)
names(df)<-c("Actual Index","Forecasted Index")
df

perc_error(df$`Actual Index`,df$`Forecasted Index`)



#Forecasted Values From ets
forecastedval2=forecast(fit2,h=109,method ='ets')
plot(forecastedval2)
#forecastedval2

#Percentage Error
df2<-data.frame(test$Date,test$Close,forecastedval2$mean)
names(df2)<-c("Date","Actual Index","Forecasted Index")

perc_error(df2$`Actual Index`,df2$`Forecasted Index`)

```





#Nifty vs Actual QNifty vs Forecasted 
```{r}

def=cbind(test$Date,df)
names(def)<-c("Date","Actual QNifty Price","Auto Arima Forecasted Qnifty Price")


finnifty=niftytest%>%select("Date","Close")
names(finnifty)<-c("Date","NiftyIndex")


#merging 
setDT(finnifty)
setDT(def)

plot1 = finnifty[def, on = c('Date','Date')]
plot1=na.omit(plot1)
plot1=plot1%>%select(-Date.1)

plot1$`ETSForecasted Qnifty Price` = forecastedval2$mean
plot1$`ARIMAForecasted Qnifty Price` = corpred

plot1

#actual vs forecasted Qnifty
ggplot(plot1, aes(Date)) + 
  geom_line(aes(y = plot1$`Actual QNifty Price`, colour = "QNifty Price")) +
  geom_line(aes(y = plot1$`ARIMAForecasted Qnifty Price`, colour = " ARIMA QNifty Price")) +
  geom_line(aes(y = plot1$`Auto Arima Forecasted Qnifty Price`, colour = "AutoArima QNifty Price")) +
  geom_line(aes(y = plot1$`ETSForecasted Qnifty Price`, colour = "ETS QNifty Price")) +
  ggtitle("Qnifty (Jan - Aug 2018) Price Prediction)")+ xlab("Date") + ylab("Price")


# normalizing the index and share price values between 0 and 1
rangenifty=range01(plot1$NiftyIndex)
rangeforcast=range01(as.numeric(plot1$`Auto Arima Forecasted Qnifty Price`))
rangeactual=range01(as.numeric(plot1$`Actual QNifty Price`))

plot2=data.frame(rangenifty,rangeforcast,rangeactual,plot1$Date)


#Nifty vs Forecasted Qnifty vs Actual Qnifty Price - Normalised
ggplot(plot2, aes(plot2$plot1.Date))+ 
  geom_line(aes(y = plot2$rangenifty, colour = "Actual Nifty")) + 
  geom_line(aes(y = plot2$rangeforcast, colour = "Forecasted Qnifty Price"))+
  geom_line(aes(y = plot2$rangeactual,colour = "Actual Qnifty Price"))+ ggtitle("Nifty vs Forecasted Qnifty vs Actual Qnifty Price - Normalised")+xlab("Date")+ ylab("Normalised Price/Index")


```

The models built generally confrom to the actual Qnifty price with an error range of 6-8%:
manual arima-8% 
auto arima-6% 
ets-6% 

But greater the range of the prediction into the future, the greater the error.

Furthermore the Indian Markets (as reflected by the Nifty) has been on a steady bull run for the past 5 years.
There were relatively few economic shocks and the positive fundametals ensured a steady trend with low volatility or reversal. Therefore it was forecasted quite accurately as this trend has continued till now.
